Implementing XBF: An Efficient Self-Describing Binary Format

### 2023-08-19 docs/presentation/presentation.tex
Update presentation

\lstdefinestyle{rust}{
	tabsize=2
}
\lstset{style=rust}

@@

\newcommand{\rot}[1]{\rotatebox{90}{#1}}

@@

\setbeameroption{show notes}

@@

\begin{frame}{Primitives}
	\begin{itemize}
		\item Boolean
		      \note[item]{True of False, one byte}
		      \note[item]{All integers are little-endian}
		      \note[item]{Integer types are equivalent to the types in stdint.h C headers}
		\item U8, U16, U32, U64, U128, U256
		      \note[item]{Unsigned Integer Types}
		      \begin{itemize}
			      \item Equivalent to uint*\_t types in C
		      \end{itemize}
		\item I8, I16, I32, I64, I128, I256
		      \note[item]{Signed Integer Types}
		      \begin{itemize}
			      \item Equivalent to int*\_t types in C
		      \end{itemize}
		\item F32, F64
		      \note[item]{Floating Point Types}
		      \begin{itemize}
			      \item IEEE 754 floating point numbers
			      \item Equivalent to float and double in C
		      \end{itemize}
		\item Bytes  (Non-UTF-8 Bytes)
		      \note[item]{String but doesn't have to be UTF-8}
		\item String (UTF-8 Bytes)
		      \note[item]{Sends length as u64 followed by UTF-8 bytes}
	\end{itemize}
\end{frame}

\begin{frame}{Primitive Metadata}
	\begin{columns}
		\begin{column}{0.5\textwidth}
			\center
			\begin{tabular}{|c|c|}
				\hline
				\textbf{Type} & \textbf{Discriminant} \\
				\hline
				Boolean       & 0                     \\
				U8            & 1                     \\
				U16           & 2                     \\
				U32           & 3                     \\
				U64           & 4                     \\
				U128          & 5                     \\
				U256          & 6                     \\
				\hline
			\end{tabular}
		\end{column}
		\begin{column}{0.5\textwidth}
			\center
			\begin{tabular}{|c|c|}
				\hline
				\textbf{Type} & \textbf{Discriminant} \\
				\hline
				I8            & 7                     \\
				I16           & 8                     \\
				I32           & 9                     \\
				I64           & 10                    \\
				I128          & 11                    \\
				I256          & 12                    \\
				F32           & 13                    \\
				F64           & 14                    \\
				Bytes         & 15                    \\
				String        & 16                    \\
				\hline
			\end{tabular}
		\end{column}
	\end{columns}
	\note[item]{Metadata is sent as a single byte discriminant value}
	\note[item]{The discriminant begin at 0 with Boolean, and increase by 1 for each type}
\end{frame}

@@

\begin{frame}{Vectors (Homogenous Lists)}
	\begin{itemize}
		\item Homogenous list of values that has a known, variable length.
		\item Length is sent as a part of the type, not in metadata.
		      \note[item]{Length is an u64, the same as String and Bytes}
		\item Length is sent, followed by the elements.
		      \note[item]{No type info about elements, that's in metadata}
		\item Inner type Information is \emph{not} sent as a part of the value.
	\end{itemize}
\end{frame}

\begin{frame}{Vector Metadata}
	\begin{itemize}
		\item Single byte discriminant value is sent (String + 1).
		\item Followed by internal type metadata (can continue recursively).
		\item Length is \emph{not} sent as a part of the metadata.
	\end{itemize}
\end{frame}

\begin{frame}[fragile]{Vector Example}
	\begin{center}
		\textbf{Vector of 16-Bit Unsigned Integers:} [42, 1024]

@@

		D16 = Discriminant of a 16-Bit Unsigned Integer

@@

\begin{frame}{Tested Formats}
	\begin{itemize}
		\item CSV
		\item MessagePack
		\item CBOR
		\item JSON
		\item XML
	\end{itemize}
\end{frame}

@@

\begin{frame}{Average Time and Bytes Read}
	\center
	\begin{tabular}{|c|c|c|}
		\hline
		\textbf{Format} & \textbf{Avg Time (ms)} & \textbf{Bytes Read} \\
		\hline
		CSV             & 18.93                  & 16,411              \\
		MessagePack     & 11.22                  & 15,565              \\
		CBOR            & 16.95                  & 25,507              \\
		JSON            & 21.91                  & 31,180              \\
		XML             & 21.87                  & 43,699              \\
		XBF             & 11.32                  & 14,686              \\
		\hline
	\end{tabular}
\end{frame}

>>>>

% pympress notes page config
\iffalse
	\setbeamertemplate{note page}
	{
		\insertslideintonotes{0.01}
		\rule{\textwidth}{0.1pt}
		%\color{blue} \scriptsize
		\insertnote
	}
	\setbeameroption{show notes on second screen}
\fi

% adjust gigantic tabs in code blocks, not actually rust format
\lstdefinestyle{rust}{
	tabsize=2
}
\lstset{style=rust}

@@

% needed this a bunch so made a macro for myself
\newcommand{\rot}[1]{\rotatebox{90}{#1}}

@@

@@

\begin{frame}{Primitives and Primitive Metadata}
	\begin{columns}
		\begin{column}{0.5\textwidth}
			\begin{itemize}
				\item Boolean
				      \note[item]{True of False, one byte}
				      \note[item]{All integers are little-endian}
				      \note[item]{Integer types are equivalent to the types in stdint.h C headers}
				\item U8, U16, U32, U64, U128, U256
				      \note[item]{Unsigned Integer Types}
				      \begin{itemize}
					      \item Equivalent to uint*\_t types in C
				      \end{itemize}
				\item I8, I16, I32, I64, I128, I256
				      \note[item]{Signed Integer Types}
				      \begin{itemize}
					      \item Equivalent to int*\_t types in C
				      \end{itemize}
				\item F32, F64
				      \note[item]{Floating Point Types}
				      \begin{itemize}
					      \item IEEE 754 floating point numbers
					      \item Equivalent to float and double in C
				      \end{itemize}
				\item Bytes  (Non-UTF-8 Bytes)
				      \note[item]{String but doesn't have to be UTF-8}
				\item String (UTF-8 Bytes)
				      \note[item]{Sends length as u64 followed by UTF-8 bytes}
			\end{itemize}
		\end{column}
		\begin{column}{0.25\textwidth}
			\center
			\begin{tabular}{|c|c|}
				\hline
				\textbf{Type} & \textbf{Value} \\
				\hline
				Boolean       & 0              \\
				U8            & 1              \\
				U16           & 2              \\
				U32           & 3              \\
				U64           & 4              \\
				U128          & 5              \\
				U256          & 6              \\
				\hline
			\end{tabular}
		\end{column}
		\begin{column}{0.25\textwidth}
			\center
			\begin{tabular}{|c|c|}
				\hline
				\textbf{Type} & \textbf{Value} \\
				\hline
				I8            & 7              \\
				I16           & 8              \\
				I32           & 9              \\
				I64           & 10             \\
				I128          & 11             \\
				I256          & 12             \\
				F32           & 13             \\
				F64           & 14             \\
				Bytes         & 15             \\
				String        & 16             \\
				\hline
			\end{tabular}
		\end{column}
	\end{columns}
\end{frame}

@@

\begin{frame}{Vectors (Homogenous Lists)}
	Values
	\begin{itemize}
		\item Homogenous list of values that has a known length.
		\item Length is sent as a part of the type, not in metadata.
		\item Inner type Information is \emph{not} sent as a part of the value.
	\end{itemize}

	Metadata
	\begin{itemize}
		\item Single byte discriminant value is sent (String + 1).
		\item Followed by internal type metadata (can continue recursively).
		\item Length is \emph{not} sent as a part of the metadata.
	\end{itemize}
\end{frame}

\begin{frame}[fragile]{Vector Example}
	\begin{center}
		Vector of 16-Bit Unsigned Integers: [42, 1024] \vfill \break

@@

		D16 = Discriminant of a 16-Bit Unsigned Integer \vfill \break

@@

\begin{frame}{Encoder and Decoder Reference Implementation}
	\begin{itemize}
		\item Written in Rust.
		      \note[item]{systems level programming language similar to C and C++, strong memory safety guarantees}
		\item Extensively Tested.
		      \note[item]{100 percent code coverage}
		      \note[item]{not going to talk a ton about the reference implementation}
		\item Majority of time was spent on this implementation.
	\end{itemize}
\end{frame}

\begin{frame}{Performance Test}
	\begin{columns}
		\begin{column}{0.5\textwidth}
			Server
			\begin{itemize}
				\item Downloads and parses 1 year of Sony stock history.
				\item 
			\end{itemize}
		\end{column}
		\begin{column}{0.5\textwidth}
				\item For each data format the client:
				      \begin{itemize}
					      \item Client sends 100 requests to the server.
					      \item Records average time to receive a response.
					      \item Records number of bytes read.
				      \end{itemize}
		\end{column}
	\end{columns}
\end{frame}

@@

\begin{frame}{Average Time and Bytes Read}
	\center
	\begin{tabular}{|c|c|c|c|c|}
		\hline
		\textbf{Format} & \textbf{Avg Time (ms)} & \textbf{Bytes Read} & \textbf{Overhead (bytes)} & \textbf{Percent Overhead} \\ 
		\hline                                                                                                                    
		CSV             & 18.93                  & 16,411              & 1,823                     & 11.1                      \\ 
		MessagePack     & 11.22                  & 15,565              & 977                       & 6.22                      \\ 
		CBOR            & 16.95                  & 25,507              & 10,919                    & 42.8                      \\ 
		JSON            & 21.91                  & 31,180              & 16,592                    & 53.2                      \\ 
		XML             & 21.87                  & 43,699              & 29,111                    & 66.6                      \\ 
		XBF             & 11.32                  & 14,686              & 98                        & 0.67                      \\ 
		\hline
	\end{tabular}
\end{frame}
